---
title: "Time Series Analysis of US New Single Family Home Sales and the Nasdaq-100 Index"
author: "STAT619 Spring 2021 - Erik Duus"
output:
  pdf_document: default
  html_document: default
theme: paper
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(astsa)
library(tseries)
library(fGarch)
library(forecast)
library(lmtest)
library(tidyverse)
library(tidyquant)  
```
### House Sales - Introduction

FRED New One Family Houses Sold: https://fred.stlouisfed.org/series/HSN1FNSA

Monthly data series NOT seasonally adjusted

Choose a 20 year window for analysis. Use ensuing 7 months for prediction error

```{r}
# New single family House sales - not seasonally-adjusted

house.df <- read.csv('./HSN1FNSA.csv', header=TRUE)
house.full_ts <- ts(data=house.df$HSN1FNSA, frequency=12, start=1963 )

# arbitrary 20-year window - avoids the GFC and covid
house.ts <- window(house.full_ts, start=1982, end=2002) 
house.future <- window(house.full_ts, start=c(2002,2), end=c(2002,8))

```
### Exploratory Data Analysis
Time series shows seasonality and trend - clearly not stationary

1 month differencing to remove trend

12 month differencing to remove seasonality

Resulting time series plot has constant mean and reasonably constant variance

```{r}
sales <- house.ts
ds <- diff(house.ts,1)
dd12s <- diff(ds, 12)
plot(house.ts, main="New House Sales (1982-2002)", ylab='', xlab='')
plot(dd12s, main='New House sales (yearly+monthly differencing)', ylab='', xlab='')

plot(cbind(sales, ds, dd12s), main='New House Sales - yearly/seasonal differencing')
acf2(dd12s, main='ACF: New House Sales (differenced)')

```
### Statistical Analysis
stationarity test show no unit root - mean is stationary

model identification: 

* seasonal ACF/PACF at lags: PACF tails off at 2 lags. ACF cuts off at 1 lag. 
likely MA(2)
* between season lags: both tail off: ARMA(1,1)

initial model: (1,1,1) / (0,1,2)(12)

* sma2 not significant

next: (1,1,1) / (0,1,1)(12)

* all coefficients significant
* ljung-box statistics are not significant

parsimonious: 

* (1,1,0) / (0,1,1)(12): bad ljung box
* (0,1,1) / (0,1,1)(12): bad ljung box



```{r}
# null hypothesis not stationary
adf.test(dd12s, k=0)
adf.test(dd12s)
pp.test(dd12s)

# null hypothesis stationary
kpss.test(dd12s)

sarima(house.ts, 1,1,1, 0,1,2,12)
sarima(house.ts, 1,1,1, 0,1,1,12) # best

sarima(house.ts, 0, 1, 1, 0,1,1,12)
sarima(house.ts, 1, 1, 0, 0,1,1,12)

# refit best model using arima() so we can use forecast()
mod1.fit = arima(house.ts, order=c(1,1,1), seasonal=list(order=c(0,1,1),period=12))
      

```
predict 7 months from end of series:

* use true data to compute RMSE
```{r}
pred1 = forecast(window(house.ts, start=1992), model=mod1.fit, h=7)
plot(pred1, main="House Predictions")
lines(house.future, lty='dashed', col='red')
legend(1992,95, legend=c('predicted','actual'), col=c('blue', 'red'), lty=1:2)

library(Metrics)
rmse(house.future, pred1$mean)
```


### Model2: Regression + ARMA errors 

FRED Leading Index for the United States https://fred.stlouisfed.org/series/USSLIND

Build regression model house ~ lead + ARMA errors
```{r}
lead.df <- read.csv('./USSLIND.csv', header=TRUE)
lead.ts <- ts(data=lead.df$USSLIND, frequency=12, start=1982)
lead.future = window(lead.ts, start=c(2002,2), end=c(2002,8))

```
### Exploratory Data Analysis

LEAD trend appears to have some correlation with HOUSE

```{r}
hsle <- ts.intersect(house = house.ts, lead = lead.ts) 

plot(hsle, main='House Sales and Lead Index Sales')
h <- decompose(hsle[,'house'])
l <- decompose(hsle[,'lead'])
htrend <- h$trend
ltrend <- l$trend
plot(cbind(htrend,ltrend), main='Trend of House Sales and Lead Index Series',
     xlab='',ylab=cbind('',''), yaxt='n')
```
\

### Statistical Analysis

fit regression house ~ lead

regression of house sales vs leading economic indicators

* coefficients are significant

examine residuals:

* trend and seasonality
* yearly and monthly difference the series
* constant mean, constant variance

check for stationarity: all tests pass. mean is stationary

variance is pretty constant

ACF analysis:

* seasonal lags - both seem to cut off at 1. would suggest ARMA(0,0)
try (1,1)

* between-season lags: tail off ARMA(1,1)


```{r}
summary(mod2.lm <-lm(house ~ lead, data=hsle))

res = resid(mod2.lm)
dd12res = diff(diff(res,12),1)

layout(matrix(c(1,1,1,1,
                2,2,2,2),
              nrow=2, byrow=TRUE))
plot(res, type="l", main='Residuals', ylab='')
plot(dd12res, type="l", main='Residuals - differenced', ylab='')

# null hypothesis not stationary
adf.test(dd12res, k=0)
adf.test(dd12res)
pp.test(dd12res)

# null hypothesis stationar
kpss.test(dd12res)

acf2(dd12res)
sarima(hsle[,'house'], 1,1,1,0,1,1,12,xreg=hsle[,'lead']) 


```


```{r}
mod2.fit = arima(hsle[,'house'],xreg=hsle[,'lead'],order=c(1,1,1), seasonal=list(order=c(0,1,1),period=12)) 
summary(mod2.fit)
coeftest(mod2.fit)
```
```{r}


mod2.fit2 <- Arima(window(hsle[,1], start=1992), 
                   model=mod2.fit, 
                   xreg=window(hsle[,2], start=1992))


pred2 = forecast(mod2.fit2, xreg=lead.future)

plot(pred2, main="House Sale Predictions - regression+ARMA")
lines(house.future, lty='dashed', col='red')
legend(1992,95, legend=c('predicted','actual'), col=c('blue', 'red'), lty=1:2)

rmse(house.future, pred2$mean)

```



