---
title: "STAT619 project Analysis 2 - Nasdaq-100/QQQ Pricing"
author: "Erik Duus"
date: "5/5/2021"
output:
  pdf_document: default
  html_document: default
---

Retrieve data from Yahoo finance: 

* use simple interpolation to approximate missing values.
* save as a local file so download not required each time
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(astsa)
library(tseries)
library(fGarch)
library(forecast)
library(lmtest)
library(tidyverse)
library(tidyquant)  

```
```{r}
options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)

# get data from YHOO and save to disk

qqq <- getSymbols("QQQ", auto.assign = FALSE)

# which(is.na(btc$'BTC-USD.Close'))
qqq <- na.approx(qqq)

#saveRDS(qqq, './qqq.rds')

```

Load local copy of pricing data

* Examine time series; it is growing exponentially
* Log transform and difference to convert to a returns series
* Mean looks stationary, variance is not constant
* Returns show volatility - GARCH may be needed
* Extract the 2016-2017 window for analysis
* Plot ACF/PACF of returns and squared returns; squared returns confirm GARCH
```{r}

# load saved Yahoo data; use adjusted closing prices
qqqall = readRDS('./qqq.rds')
qqqall <- qqqall$QQQ.Adjusted

qqqr <- diff(log(qqqall))
plot(qqqr, main='QQQ daily returns')


qqq <- window(qqqall, start='2016-01-01',end='2018-01-01')
qqqr <- window(qqqr, start='2016-01-01',end='2018-01-01')

layout(matrix(c(1,1,1,1,
                2,2,2,2),
              nrow=2, byrow=TRUE))

plot(qqqall, main='QQQ closing price')
plot(qqqr, main='QQQ daily returns 2016-2017')

acf2(qqqr, main='QQQ returns')
acf2(qqqr^2, main='QQQ returns squared')
```
\ confirm stationarity of mean through unit root tests
```{r}
# null hypothesis not stationary
adf.test(qqqr, k=0)
adf.test(qqqr)
pp.test(qqqr)

# null hypothesis stationary
kpss.test(qqqr)
```

\

Run auto.arima to see what it suggests;diagnostics of fitted model also show
variance in residuals, confirming garch.

```{r}
auto.arima(qqqr, seasonal=FALSE)
sarima(qqqr, 1,0,0)
```
 \
 
 ACF and PACF of series shows a tiny bit of auto-correlation, suggesting 
 ARMA(1,1)
 
 ACF and PACF of squared series shows both decaying, suggesting GARCH(1,1)
 
 Fit GARCH(1,1) to start as a baseline
 
```{r}
gf <- garchFit(~garch(1,1), data=qqqr, cond.dist='std', trace=FALSE)
summary(gf)

layout(matrix(c(1,1,1,1,
                1,1,1,1,
                2,2,4,4,
                2,2,4,4,
                3,3,4,4,
                3,3,4,4),nrow=6, byrow=TRUE))
plot(gf, which=9)
plot(gf, which=10)
plot(gf, which=11)
plot(gf, which=13)


```
\ 

model coefficients are significant, but QQ-plot is not that normal. perhaps 
ARMA features are required.

several iterations get us to ARMA(5,3)+GARCH(1,1)

** (NOTE ARMA(5,3) is best model but wont predict correctly) **
```{r}
gf53 <- garchFit(~arma(5,3)+garch(1,1), data=qqqr ,cond.dist='std', trace=FALSE)
summary(gf53)

layout(matrix(c(1,1,1,1,
                1,1,1,1,
                2,2,4,4,
                2,2,4,4,
                3,3,4,4,
                3,3,4,4),nrow=6, byrow=TRUE))
plot(gf53, which=9)
plot(gf53, which=10)
plot(gf53, which=11)
plot(gf53, which=13)


```
\ 

QQ plot is much better. residual plot and ACF plots suggest residuals are white noise.
Ljung-Box tests are also suggestive of white noise residuals.

Conclude this model is accurate and use for predictions.

#### NOTE: fgarch package produces errors on this model when forecasting!
revert to ARMA(2,2)+GARCH(1,1) for forecast

```{r}
gf1 <- garchFit(~arma(2,2)+garch(1,1), data=qqqr ,cond.dist='std', trace=FALSE)
summary(gf1)

layout(matrix(c(1,1,1,1,
                1,1,1,1,
                2,2,4,4,
                2,2,4,4,
                3,3,4,4,
                3,3,4,4),nrow=6, byrow=TRUE))
plot(gf1, which=9)
plot(gf1, which=10)
plot(gf1, which=11)
plot(gf1, which=13)


```
\

predict 14 days of future returns. recall this is a daily returns series, so
these are predictions of future daily returns.

```{r}
### 14 day predictions of returns - PLOT
preds <- predict(gf1, n.ahead=14, plot=TRUE)
preds
```
\ 

to produce pricing predictions the returns series needs to be applied
to the tail of the price series, and then undifferenced and exp() applied.


```{r}
# convert forecast means and intervals to prices
mf <- diffinv(preds$meanForecast, xi=log(qqq[503]))
mf <- mf[2:15] # drop first seed value (from original series)
li <- mf * (1 + preds$lowerInterval)
ui <- mf * (1 + preds$upperInterval)
mf <- exp(mf)
li <- exp(li)
ui <- exp(ui)

# get original time series plus the actual values for prediction interval
# create prediction, upper and lower interval series aligned with actuals
qqq_true <- window(qqqall, start='2017-07-01',end='2018-01-22')
qqq_mf <- tail(qqq_true, 14)
qqq_mf[,1] <- mf
qqq_li <- tail(qqq_true, 14)
qqq_li[,1] <- li
qqq_ui <- tail(qqq_true, 14)
qqq_ui[,1] <- ui


plt <- plot(ylim=c(130,175), qqq_true, type='l', main='QQQ predictions - ARMA(1,1)+GARCH(5,3)', ylab='Price', xlab='')
plt <- lines(qqq_mf, lty=2, lwd=1, col='blue')
plt <- lines(qqq_li, lty=4, lwd=1, col='red')
plt <- lines(qqq_ui, lty=4, lwd=1, col='red')
plt <- addLegend('topleft', legend.names=c('predicted','lower', 'upper'),
          lty=c(2,4,4), lwd=c(1,1,1),col=c('blue', 'red', 'red'))
plt


```

\ 

compute RMSE of predictions
```{r}
library(Metrics)
rmse(tail(qqq_true,14), tail(qqq_mf,14))
```


